{Exciton-polariton Josephson junctions at finite temperatures}

### 2017-03-15 writing/phase_transition/phase_transition.tex
phase_transition: revision, nice potential picture

\author{A. V. Kavokin}
\affiliation{Spin Optics Laboratory, St. Petersburg State University, Ul’anovskaya, Peterhof, St. Petersburg 198504, Russia}
\affiliation{School of Physics and Astronomy, University of Southampton, SO17 1BJ Southampton, United Kingdom}

@@

\begin{abstract}
We consider the problem of achievement of quantum regime for relative phase of trapped in W -- potential low branch exciton polariton condensates taken at finite temperatures and finite polariton lifetime.
Non-standard Bose-Hubbard model for an exciton polariton Josephson junction (EPJJ) that is characterised by complex potential energy landscapes (PEL) consisting of sets of barriers and wells is proposed.
We have shown that the classical (thermal) activation and quantum tunneling for the phase represent two regimes depending on temperature, polariton condensate density and distance between the junctions.
The transition between these two regimes exhibits universal features of first or second order phase transition (PT) depending on the PEL for the relative phase of two polariton condensates.
In the presence of dissipation quantum system exhibits the non-equilibrium PT from quantum {\blue (where tunneling is possible)} to classical regime, where the tunelling is efficient and where it is suppressed, respectively.
The obtained results are discussed in details in the framework of the quantum annealing problem with exciton polariton condensates.

@@

In the last decade investigations of exciton polariton Bose-Einstein condensates (BEC) in various type of semiconductor microstructures represented an important mainstream of current photonic and semiconductor studies, see e.g. \cite{Sanvitto, Guillet}.
The microcavity exciton polaritons are bosonic quasiparticles representing  admixtures of the quantized cavity mode and quantum well excitons.
Semiconductor microcavities are promising for various optoelectronic applications where the quantum  matter-field interface plays crucial role.
In particular, it is worth to mention demonstrations of a polariton laser with electrical injection pump \cite{Bhattacharya,Schneider}, amplifiers \cite{Niemietz}, switches \cite{Amo_2010}, transistors \cite{Ballarini}, polariton circuits and optical logic elements \cite{Sturm,Liew}.
Although exciton polariton condensates exhibit Bose-Einstein statistics  above the lasing threshold under the phase transition (PT) condition and macroscopic occupation of the ground state at certain pumping rate that is less than for convenient lasers, they are not in true thermodynamic equilibrium state \cite{Byrnes_2014,Sun}. Non-equilibrium features of exciton polariton condensates play important role in various manifestations of their collective (many body) states such as superfluidity \cite{Carusotto_2013,Amo_2009}, quantized vortices \cite{Lagoudakis_2008,Lagoudakis_2009}, soliton formation \cite{Sich}, Josephson oscillations and macroscopic self-trapping \cite{Abbarchi,Lagoudakis_2010}.

Recently Yongbao Sun et al, in Ref. \cite{Snoke_2017} reported observation of a quasi-equilibrium low branch exciton polariton condensate in a high Q-factor microcavity possessing very a comparatively very long photon lifetime of $135ps$. Notably, exciton polaritons with long lifetime are promising candidates for quantum technologies, cf. \cite{Demirchyan}. However, since exciton polariton BEC's in present experiments are observed at high enough (up to the room) temperatures the quantum properties of such a states should be examined even at full thermal equilibrium.

The problem of distinguishability of statistically classical and quantum regimes for exciton polariton condensates behavior is very actual for their practical applications and especially for their possible applications in quantum information technologies, see e.g. \cite{Dominici}.
Fast switching properties (the typical switching time of a few picoseconds), relatively strong nonlinear response and flexibility to external optical and/or electrical pump, spin degrees of freedom made microcavity polaritons potentially very promising for quantum computation and quantum information processing objectives, see, e.g. \cite{Demirchyan,Pagel,Kyriienko,Solnyshkov_2015, Dominici}.

Especially it is important to stress here quantum annealing problem physically based on the searching algorithm for the global minimum of a potential energy landscape (PEL) consisting of a set of barriers and wells, see e.g.  \cite{Santoro, Das}. In purely classical  (thermal) regime the {bosonic quasipartcles cross barriers stochastically} at finite temperature with the help of thermal activation process if the thermal energy is enough. Contrary, in a quantum limit the same system undergoes quantum tunneling through the barrier. Obviously, the shape of the barrier plays an essential role in this case, cf. \cite{Das, Lewenstein}. It has been shown recently in \cite{Lewenstein} that the annealing algorithm relies in general case on combination of thermal annealing and quantum tunneling. The     
collective (bosonic) character of polariton condensation could be used for the acceleration of the physical implementation of the search algorithm \cite{Yan}.

In the paper we consider a generic problem of quantum-classical phase transition (PT) with exciton polariton condensates formed in a semiconductor microcavity in the presence of Josephson effect, cf. \cite{Chudnovsky_1997,Aleiner, Shelykh_2008, Borgh_2010}. To be more specific we examine simply a two junction model.

In general this problem relates to dissipative tunneling problem or, tunneling at finite (non-zero) temperatures, cf. \cite{Caldeira, Larkin, Riseborough}.
Traditionally,  studies in this field are limited by consideration of superconductor devices \cite{Ankerhold} where macroscopic quantum tunneling phenomena plays essential role.
In particular, it is worth to mention here the Schr\"odinger cat state formation \cite{Leggett} and the 	design of quantum gates with superconductor qubits for quantum computing where a macroscopic quantum coherence for the phase is important \cite{Makhlin}.
Recently the problem of quantum tunneling and quantum-classical PT problem have been examined for other two-level (spin) systems, like atomic BEC's \cite{Zhang}, and small ferromagnetic particles, \cite{Owerre}.

In the paper we consider an \textit{extended model} for exciton polariton Josephson junctions (EPJJs) that relates to so-called non-standard Bose-Hubbard models (see \cite{Dutta} and references therein) when energy of polariton-polariton scattering contributes into the tunneling parameter matrix elements, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010, Solnyshkov_2008, Sarchi}.
In particularly, we pay attention to temperature dependent quantum critical phenomena occurring in the presence of macroscopic tunneling.
We aimed at formulating necessary criteria to obtain quantum phase regimes with trapped exciton polaritons suitable for realisation of quantum annealing purposes and creation of exciton polariton qubit gates.

This paper is arranged as follows.
In Sec. \ref{sec:quantum_phase} we introduce the quantum phase model for extended model of EPJJ of the coupled condensates and discuss various regimes for it at zero temperature and infinite exciton polariton lifetime. Section \ref{sec:quantum_classical} establishes  results obtained for quantum-classical PT problem and considered for quantum phase  at finite temperatures.  The influence of non-equilibrium effects with microcavity exciton polaritons possessing finite lifetime is discussed in \ref{sec:non-equilibrium}.      
In Conclusion we summarize the results obtained.
In Appendix \ref{sec:model} we explain in details  extended model of EPJJ within the tight-binding approach. 
\section{The Quantum Phase Model \label{sec:quantum_phase}}

We start with the Hamiltonain 
%
\begin{equation}
\hat{H} = \alpha \hat{S}_z^2 + \beta \hat{S}_x^2 - B \hat{S}_x.
\label{eq:hamiltonian_spin}
\end{equation}
that characterize weakly interacting exciton polariton condensate Josephson junction device in terms of pseudo-spin operators $\hat{S}_i$, $i=x,y,z$.
The parameters $\alpha$, $\beta$ and $B$ reflect  Josephson junction properties and determined in Appendix \ref{sec:model}. 
Here we propose quantum mechanical approach to the EPJJ  phase recasting pseudo-spin operators in  (\ref{eq:hamiltonian_spin}) through the differential representation defined as
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & -i\dfrac{d}{d \phi}
\end{align}
\label{eq:pseudo_spin_diff}
\end{subequations}
%
where  $\phi$ is the phase variable.
The operators determined in Eq. (\ref{eq:pseudo_spin_diff}) in \eqref{eq:hamiltonian_spin}
 obey familiar $SU(2)$ algebra commutation relations $[\hat{S}_i, \hat{S}_j] = i\epsilon_{ijk}\hat{S}_k,  i,j,k=x,y,z$.
After straitforward calculations from (\ref{eq:hamiltonian_spin}) one can obtain
%
\begin{equation}
\begin{array}{lcl}
H & = & -(\alpha - \beta \sin^2 \phi) \dfrac{d^2}{d \phi^2} - (\beta(s-\dfrac{1}{2})\sin{2\phi} - B\sin{\phi}) \dfrac{d}{d \phi} \\
&& - Bs \cos \phi - \beta s^2 \sin^2 \phi - \beta s \cos^2 \phi.
\end{array}
\label{eq:hamiltinian_s}
\end{equation}
%
We also assume macroscopically large particle number, i.e. the $s = N/2 >> 1$ is C - number. Using (\ref{eq:hamiltinian_s})  for Schrodinger equation we obtain   
%
\begin{equation}
\begin{array}{l}
(\alpha - \beta \sin^2 \phi) \dfrac{d^2 \Phi}{d \phi^2} + (\beta s\sin{2\phi}-B\sin{\phi})) \dfrac{d \Phi}{d \phi} \\
+ (E + B s \cos \phi + \beta s^2 \sin^2 \phi) \Phi = 0,
\end{array}
\label{eq:schrodinger}
\end{equation}
%
where $\Phi \equiv \Phi(\phi)$ is unknown $2\pi$-periodic wavefunction that characterizes quantum phase properties, cf. \cite{Anglin}.
%ng that Eqs. (\ref{eq:hamiltinian_s}), (\ref{eq:schrodinger}) can be obtained by applying so-called phase-state representation, {\red cf. [  ]}.
%
%\begin{equation}
%\ket{\psi} = \dfrac{1}{2 \pi} \int_{-\infty}^{+\infty} d \phi \Phi(\phi) \ket{\phi}
%\end{equation}
%
%where
%
%\begin{equation}
%\ket{\phi} = \dfrac{1}{\sqrt{N!}} \Big[ e^{i\phi/2} \psi_1^\dagger + e^{-i \phi/2} \psi_2^\dag \Big]^N \ket{0}_1\ket{0}
%\end{equation}
%
%is macroscopic coherent spin-state (macroscopic qubit state) that plays important role in current quantum information protocols operating with $N$-particle condensates, cf. \cite{Byrnes_2012}.
It is possible to eliminate the term with first derivative in Eq. (\ref{eq:schrodinger}) applying the transformation
%
\begin{equation}
\Phi(\phi) = \Psi(z) \exp \Big( s \ln \dn{z} - \dfrac{\Lambda s}{2 \sqrt{\lambda (1 - \lambda)}} \arctan \Big( \sqrt{\dfrac{\lambda}{1 - \lambda}} \cn{z} \Big) \Big),
\label{eq:transformation_of_phi}
\end{equation}
%
where $z = \int \limits_0^\phi \dfrac{d \xi}{\sqrt{1 - \lambda \sin^2 \xi}} = F(\phi, \lambda)$ is new phase variable; $F(\phi, \lambda)$ is incomplete elliptic integral of the first kind.
In (\ref{eq:transformation_of_phi}) we have introduced dimension-less parameters $\Lambda = \dfrac{G}{\alpha s}$, $\lambda = \dfrac{\beta}{\alpha}$ playing crucial role for our model.
Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we arrive to Schr\"odinger-like equation
% 
\begin{equation}
\alpha \frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0
\label{eq:schrodinger_usual}
\end{equation}
%
characterizing quantum particle with the ``mass''  $m = \dfrac{\hbar^2}{2 \alpha}$ and energy $E$, that moves in the potential $V(z) = \alpha s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}.
\label{eq:potential}
\end{equation}
%

@@

%
\begin{figure}[ht]
\begin{center}\includegraphics[width=1\linewidth]{pic/potentials.eps}
\end{center}
\caption{
Effective dimension-less potential $V_0(z)$ for (a) $\Lambda = 100$, (b) $\Lambda = 0.5$ , and (c) $\Lambda = 0.01$ as a function of elliptic integral phase coordinate $z$.
Points $\pm z_{min}$ and $\pm z_{max}$ in (b) relay to global minima and maxima of the potential, correspondingly, occurring with periodicity $4K(\lambda)$. \label{phase_potential}}
\end{figure}
%

Analysis of the quantum phase can be performed  in three domains determined by vital $\Lambda$--parameter values, cf. \cite{Anglin}.

@@

To  be more specific we consider the tunneling in  effective $W$ -- like potential represented in Fig. \ref{phase_potential}b  inherent to  Josephson regime. We examine thermodynamically equilibrium (with infinitely large  exciton polariton lifetime) system of exciton polaritons at finite temperature $T$.  Noticing, that in  recent experiments  described in Ref. \cite{Snoke_2017} polariton lifetime attains  $275ps$ at resonance that seems to be enough to justify the quantum processes described below.

Transition between two stable states (say, between points $-z_{min}$ and $+z_{min}$ in Fig. \ref{phase_potential}b) can happen  either through the quantum tunneling or, in classical way, due to the thermal activation. 
Obviously, at high temperatures such as $k_{B}T\ge\Delta V$ ($\Delta V$ is height of the barrier between two states with minimum of potential energy) the particle jumps over the barrier governed by familiar thermoactivation (Arrhenius) escape rate $\Gamma \sim e^{-\Delta V /k_{B}T}$, cf. \cite{Larkin}. This process inherent to classical (thermal) annealing problem \cite{Das}.

In the ``low temperature'' limit $k_{B}T\ll\Delta V$ particle undergoes quantum tunneling through the barrier with vanishing rate, that is important especially for quantum annealing scheme. 
%with the rate $\Gamma \sim e^{-B/\hbar}$, where $B$ is the instanton action.
%Now we elucidate conditions for their realization for exciton polariton Josephson junction problem.

@@

Let us briefly discuss observation feasibility of quantum-classical PT for exciton polariton BEC's obtained in experiments. In particular, for experimentally accessible polariton interaction strength $\alpha N=0.6meV$, given for narrow-band semiconductor samples, the  $T_{0}$ is about $1.2K$ that is comparable with typical temperature of condensation, or, with temperature of Berezinsky-Kosterlitz-Thouless PT considered for dilute weakly interacting exciton polariton gas being at thermal equilibrium,cf. \cite{Sanvitto,Guillet}. Practically some enhancement of $T_{0}$ is possible by increasing of exciton polariton density and by improvement of experiment conditions, cf. \cite{Snoke_2017}.  
%
%\begin{equation}
%T_{c}^{(2)} = (T_{0}/ 2) %\sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}.
%\label{eq:second_order_2}
%\end{equation}
%

@@

In order,  characteristic  temperature $T_{0}=1.2K$ corresponds to thermon effective lifetime  $\tau_{0}=3.5ps$ for estimations given above. Here we examine situation when exciton polariton lifetime $\tau_{pol}$ is comparable with $\tau_{0}$. Namely  this situation is realized in most of current experiments with non-equilibrium exciton polariton condensates, see e.g. [   ].

>>>>

\author{A. V. Kavokin}
\affiliation{Spin Optics Laboratory, St. Petersburg State University, Ul’anovskaya, Peterhof, St. Petersburg 198504, Russia}
\affiliation{School of Physics and Astronomy, University of Southampton, SO17 1BJ Southampton, United Kingdom}
\affiliation{Istituto CNR-SPIN, Viale del Politecnico 1, I-00133, Rome, Italy}

@@

\begin{abstract}
We consider finite temperature effects in a non-standard Bose-Hubbard model for an exciton polariton Josephson junction (EPJJ) that is characterised by a complex potential energy landscapes (PEL) consisting of sets of barriers and wells.
We have shown that the transition between  thermal  activation (classical) and tunneling (quantum) regimes exhibits universal features of the first and second order phase transition (PT) depending on the PEL for two polariton condensates.
In the presence of dissipation the relative phase of two condensates exhibits non-equilibrium PT from quantum regime characterized by efficient tunneling of polaritons to the regime of permanent Josephson or Rabi oscillations, where the tunneling is suppressed, respectively.
The obtained results are discussed in details in the framework of the quantum annealing problem with exciton polariton condensates.

@@

In the last decade investigations of exciton polariton Bose-Einstein condensates (BEC) in various type of semiconductor microstructures became and important area of research in photonics and semiconductor physics, see e.g. \cite{Sanvitto, Guillet}.
Microcavity exciton polaritons are bosonic quasiparticles representing admixtures of quantized cavity photons mode and quantum well excitons.
Semiconductor microcavities are promising for various optoelectronic applications where the quantum  matter-field interface plays crucial role.
In particular, it is worth to mention experimental demonstrations of a polariton laser with electrical injection pump \cite{Bhattacharya,Schneider}, amplifiers \cite{Niemietz}, switches \cite{Amo_2010}, transistors \cite{Ballarini}, polariton circuits and optical logic elements \cite{Sturm,Liew}.
Although exciton polariton condensates exhibit the Bose-Einstein statistics above the lasing threshold and are characterized by a macroscopic occupation of the ground state at certain pumping rate that is less than the threshold pumping for convenient lasers, they are not in a true thermodynamic equilibrium state \cite{Byrnes_2014,Sun}. Non-equilibrium features of exciton polariton condensates play an important role in various manifestations of their collective (many body) states such as superfluidity \cite{Carusotto_2013,Amo_2009}, quantized vortices \cite{Lagoudakis_2008,Lagoudakis_2009}, soliton formation \cite{Sich}, Josephson oscillations and macroscopic self-trapping \cite{Abbarchi,Lagoudakis_2010}.

Recently Yongbao Sun et al, in Ref. \cite{Snoke_2017} reported the observation of a quasi-equilibrium low branch exciton polariton condensate in a high Q-factor microcavity possessing a comparatively very long photon lifetime of $135ps$. Notably, exciton polaritons with long lifetime are promising candidates for quantum technologies, cf. \cite{Demirchyan}. However, since exciton polariton BEC's in present experiments are observed at high enough temperatures the quantum properties of such a states should be examined even at full thermal equilibrium.

The problem of distinguishability of statistically classical and quantum regimes for exciton polariton condensates is very important for their possible applications in quantum information technologies, see e.g. \cite{Dominici}.
Fast switching properties (the typical switching time of a few picoseconds), relatively strong nonlinear response and flexibility to external optical and/or electrical pump, spin degrees of freedom made microcavity polaritons potentially very promising for quantum computation and quantum information processing objectives, see, e.g. \cite{Demirchyan,Pagel,Kyriienko,Solnyshkov_2015, Dominici}.

Especially, it is important to emphasize here quantum annealing problem physically based on the searching algorithm for the global minimum of a potential energy landscape (PEL) consisting of a set of barriers and wells, see e.g.  \cite{Santoro, Das}.
In a purely classical  (thermal) regime the {bosonic quasipartcles cross barriers stochastically} at finite temperature with the help of thermal activation process if the thermal energy is large enough. Contrary, in a quantum limit the same system undergoes quantum tunneling through the barrier. Obviously, the shape of the barrier plays an essential role in this case, cf. \cite{Das, Lewenstein}. It has been shown recently in \cite{Lewenstein} that the annealing algorithm relies in general case on combination of thermal annealing and quantum tunneling. The     
collective (bosonic) character of polariton condensation could be used for the acceleration of the physical implementation of the search algorithm \cite{Yan}.

In this paper we consider theoretically the finite temperature effects for exciton polariton condensates formed in a semiconductor microcavity in the presence of the Josephson effect, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010}. 
In particular, we are interesting in generic problem of quantum-classical phase transition (PT) explicitly examined for a two junction model \cite{Chudnovsky_1997}.

In general, this problem is related to the dissipative tunneling problem or tunneling at finite (non-zero) temperatures, cf. \cite{Caldeira, Larkin, Riseborough}.
Traditionally,  studies in this field are limited by consideration of superconductor devices \cite{Ankerhold} where macroscopic quantum tunneling phenomena plays an essential role.
In particular, it is worth to mention here the Schr\"odinger cat state formation \cite{Leggett} and the 	design of quantum gates with superconductor qubits for quantum computing where a macroscopic quantum coherence for the phase is important \cite{Makhlin}.
Recently, the problem of quantum tunneling and quantum-classical PT problem have been examined for other two-level (spin) systems, like atomic BEC's \cite{Zhang}, and small ferromagnetic particles, \cite{Owerre}.

In the paper we consider an \textit{extended model} for exciton polariton Josephson junctions (EPJJs) that relates to so-called non-standard Bose-Hubbard models (see \cite{Dutta} and references therein) when energy of polariton-polariton scattering contributes to the tunneling parameter matrix elements, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010, Solnyshkov_2008, Sarchi}.
In particularly, we pay attention to temperature dependent quantum critical phenomena occurring in the presence of macroscopic tunneling.
We aimed at formulating necessary criteria to obtain quantum phase regimes with trapped exciton polaritons suitable for realisation of quantum annealing algorythms and creation of exciton polariton qubit gates.

This paper is organised as follows.
In Sec. \ref{sec:quantum_phase} we introduce the quantum phase model for extended model of EPJJ of the coupled condensates and discuss various regimes for it at zero temperature and infinite exciton polariton lifetime.
Section \ref{sec:quantum_classical} presents the results obtained for a quantum-classical PT problem and addresses the quantum phase at finite temperatures.
The influence of non-equilibrium effects with microcavity exciton polaritons possessing a finite lifetime is discussed in \ref{sec:non-equilibrium}.      
In Conclusion, we summarize the obtained results.
In the Appendix \ref{sec:model} we explain in details  extended model of EPJJ within the tight-binding approximation.

\section{The Quantum Phase Model \label{sec:quantum_phase}}

{\red

We consider a system of two exciton-polariton condensates localised in lateral potential traps in a planar semiconductor microcavity.
The condensates are separated by a potential barrier $U(x)$.
Polaritons are pumped to the condensates from a reservoir of incoherent excitons.
They decay radiatively.
We account for the repulsive intereaction between exciton-polaritons.
The finite temperature of the polariton gas is accounted for in the thermal equilibrium approximation.
We use two-mode representation of the $\hat{\psi}$ - operator that describe the behaviour of the condensates:
%
\begin{equation}
\hat{\psi} = \hat{\psi}_1(t) \Phi_1(x) + \hat{\psi}_2(t) \Phi_2(x).
\end{equation}
%
Here $\Phi_{1,2} = \dfrac{\Phi_+ \pm \Phi_-}{\sqrt{2}}$ where $\Phi_{\pm}(x)$ are symmetric and antisymmetric real wavefuntions obeying the stationary Gross-Pitaevsky equation $\mu_{\pm} \Phi_{\pm} = -\dfrac{\hbar^2}{2m_{pol}} \dfrac{d^2 \Phi_{\pm}}{dx^2} + U(x) \Phi_{\pm} + g \Phi_{\pm}^3$.

%
\begin{figure}[ht]
\center{\includegraphics[width=0.5\linewidth]{pic/potential_picture.eps}}
\caption{ \dots \label{pic:potential_picture}}
\end{figure}
%

We can express the Hamiltonian of such systems in terms of pseudo-spin operators as follows:
%
\begin{equation}
\hat{H} = \alpha \hat{S}_z^2 + \beta \hat{S}_x^2 - B \hat{S}_x
\label{eq:hamiltonian_spin}
\end{equation}
Is characterizes a weakly interacting exciton polariton condensate Josephson junction device.
The parameters $\alpha = 2 \gamma_{+-}$, $\beta = \dfrac{\gamma_{++} + \gamma_{--} - 2 \gamma_{+-}}{2}$ and $B = \mu_- - \mu_+$, $\gamma_{ij} = g \int \Phi_i^2 \Phi_j^2 dx, i,j \in \{+,-\}$ reflect Josephson junction properties, see Appendix \ref{sec:model} for details.
}
Here we propose quantum mechanical approach to the EPJJ  phase recasting pseudo-spin operators in  (\ref{eq:hamiltonian_spin}) through the differential representation defined as
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & -i\dfrac{d}{d \phi}
\end{align}
\label{eq:pseudo_spin_diff}
\end{subequations}
%
where  $\phi$ is the phase variable.
The operators determined in Eq. (\ref{eq:pseudo_spin_diff}) in \eqref{eq:hamiltonian_spin}
 obey familiar $SU(2)$ algebra commutation relations $[\hat{S}_i, \hat{S}_j] = i\epsilon_{ijk}\hat{S}_k,  i,j,k=x,y,z$.
After straightforward calculations from (\ref{eq:hamiltonian_spin}) one can obtain
%
\begin{equation}
\begin{array}{lcl}
H & = & -(\alpha - \beta \sin^2 \phi) \dfrac{d^2}{d \phi^2} - (\beta(s-\dfrac{1}{2})\sin{2\phi} - B\sin{\phi}) \dfrac{d}{d \phi} \\
&& - Bs \cos \phi - \beta s^2 \sin^2 \phi - \beta s \cos^2 \phi.
\end{array}
\label{eq:hamiltinian_s}
\end{equation}
%
We also assume macroscopically large number of particles, i.e. the $s = N/2 >> 1$ is C - number. Using (\ref{eq:hamiltinian_s})  for Schrodinger equation we obtain   
%
\begin{equation}
\begin{array}{l}
(\alpha - \beta \sin^2 \phi) \dfrac{d^2 \Phi}{d \phi^2} + (\beta s\sin{2\phi}-B\sin{\phi})) \dfrac{d \Phi}{d \phi} \\
+ (E + B s \cos \phi + \beta s^2 \sin^2 \phi) \Phi = 0,
\end{array}
\label{eq:schrodinger}
\end{equation}
%
where $\Phi \equiv \Phi(\phi)$ is unknown $2\pi$-periodic wavefunction that characterizes quantum phase properties, cf. \cite{Anglin}.
%ng that Eqs. (\ref{eq:hamiltinian_s}), (\ref{eq:schrodinger}) can be obtained by applying so-called phase-state representation, {\red cf. [  ]}.
%
%\begin{equation}
%\ket{\psi} = \dfrac{1}{2 \pi} \int_{-\infty}^{+\infty} d \phi \Phi(\phi) \ket{\phi}
%\end{equation}
%
%where
%
%\begin{equation}
%\ket{\phi} = \dfrac{1}{\sqrt{N!}} \Big[ e^{i\phi/2} \psi_1^\dagger + e^{-i \phi/2} \psi_2^\dag \Big]^N \ket{0}_1\ket{0}
%\end{equation}
%
%is macroscopic coherent spin-state (macroscopic qubit state) that plays important role in current quantum information protocols operating with $N$-particle condensates, cf. \cite{Byrnes_2012}.
It is possible to eliminate the term with first derivative in Eq. (\ref{eq:schrodinger}) applying the transformation
%
\begin{equation}
\Phi(\phi) = \Psi(z) \exp \Big( s \ln \dn{z} - \dfrac{\Lambda s}{2 \sqrt{\lambda (1 - \lambda)}} \arctan \Big( \sqrt{\dfrac{\lambda}{1 - \lambda}} \cn{z} \Big) \Big),
\label{eq:transformation_of_phi}
\end{equation}
%
where $z = \int \limits_0^\phi \dfrac{d \xi}{\sqrt{1 - \lambda \sin^2 \xi}} = F(\phi, \lambda)$ is new phase variable; $F(\phi, \lambda)$ is incomplete elliptic integral of the first kind.
In (\ref{eq:transformation_of_phi}) we have introduced dimension-less parameters $\Lambda = \dfrac{B}{\alpha s}$, $\lambda = \dfrac{\beta}{\alpha}$ playing crucial role for our model.
Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we arrive to Schr\"odinger-like equation
% 
\begin{equation}
\alpha \frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0
\label{eq:schrodinger_usual}
\end{equation}
%
characterizing quantum particle with the ``mass''  $m = \dfrac{\hbar^2}{2 \alpha}$ and energy $E$, that moves in the potential $V(z) = \alpha s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}.
\label{eq:potential}
\end{equation}
%

@@

%
\begin{figure}[ht]
\begin{center}\includegraphics[width=1\linewidth]{pic/potentials.eps}
\end{center}
\caption{
Effective dimension-less potential $V_0(z)$ for (a) $\Lambda = 100$, (b) $\Lambda = 0.5$ , and (c) $\Lambda = 0.01$ as a function of elliptic integral phase coordinate $z$.
Points $\pm z_{min}$ and $\pm z_{max}$ in (b) relay to global minima and maxima of the potential, correspondingly, occurring with periodicity $4K(\lambda)$. \label{phase_potential}}
\end{figure}
%
Analysis of the quantum phase can be performed  in three domains determined by vital $\Lambda$--parameter values, cf. \cite{Anglin}.

@@

To  be more specific we consider the tunneling in  effective $W$ -- like potential represented in Fig. \ref{phase_potential}b  inherent to  Josephson regime. We examine thermodynamically equilibrium  system of exciton polaritons at finite temperature $T$ that is the system with infinitely large  exciton polariton lifetime.  Noticing, that in  recent experiments  described in Ref. \cite{Snoke_2017} polariton lifetime attains  $275ps$ at resonance that seems to be enough to justify the equilibrium  processes described below.

Transition between two stable states (say, between points $-z_{min}$ and $+z_{min}$ in Fig. \ref{phase_potential}b) can happen  either through the quantum tunneling or, in classical way, due to the thermal activation. 
Obviously, at high temperatures such as $k_{B}T\ge\Delta V$ ($\Delta V$ is height of the barrier between two states with minimum of potential energy) the particle jumps over the barrier governed by familiar thermoactivation (Arrhenius) escape rate $\Gamma \sim e^{-\Delta V /k_{B}T}$, cf. \cite{Larkin}. This process inherent to classical (thermal) annealing problem \cite{Das}. 
In the ``low temperature'' limit $k_{B}T\ll\Delta V$ particle undergoes quantum tunneling through the barrier with vanishing rate, that is important especially for quantum annealing scheme. 
%with the rate $\Gamma \sim e^{-B/\hbar}$, where $B$ is the instanton action.
%Now we elucidate conditions for their realization for exciton polariton Josephson junction problem.

@@

Let us briefly discuss observation feasibility of quantum-classical PT for exciton polariton BEC's obtained in experiments. In particular, for experimentally accessible polariton interaction strength $\alpha N=0.6meV$, given for narrow-band semiconductor samples, the  $T_{0}$ is about $1.2K$ that corresponds to to thermon effective lifetime  $\tau_{0}=3.5ps$. The value of $T_{0}$ is comparable with typical temperature of condensation, or, with temperature of Berezinsky-Kosterlitz-Thouless PT considered for dilute weakly interacting exciton polariton gas being at thermal equilibrium,cf. \cite{Sanvitto,Guillet}. Practically some enhancement of $T_{0}$ is possible by increasing of exciton polariton density and by improvement of experiment conditions, cf. \cite{Snoke_2017}.  
%
%\begin{equation}
%T_{c}^{(2)} = (T_{0}/ 2) %\sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}.
%\label{eq:second_order_2}
%\end{equation}
%

@@

In order,  condition (\ref{eq:lifetime}) is fulfilled in [19]. Here we examine situation when exciton polariton lifetime $\tau_{pol}$ is comparable with $\tau_{0}$. Namely  this situation is realized in most of current experiments with non-equilibrium exciton polariton condensates, see e.g. [1,2, 16-18].

To be more specific we consider the influence of the dissipation to  EPJJ quantum phase properties in adiabatic limit, cf. \cite{Sols}.
We start from GP equations obtained from  (\ref{eq:hamiltonian}) for mean fields $\psi_{1,2}=<\hat\psi_{1,2}>$  at $\Gamma = 0$ ($B = G$),  which reads as 
\begin{equation}
i \hbar \dot{\psi}_{1,2} = -i \kappa \psi_{1,2} + (A|\psi_{1,2}|^2 + 2C |\psi_{2,1}|^2) \psi_{1,2} - (B/2 - C \psi_{1,2}^* \psi_{2,1}) \psi_{2,1}. 
\label{eq:GP}
\end{equation}
In (\ref{eq:GP}) we introduced dissipation  term with $\kappa \simeq \hbar/\tau_{pol}$.
Defining new variables $\Psi_{1,2}$ as  $\psi_{1,2} = \Psi_{1,2} \exp(-\kappa t / \hbar)$ for mean field pseudo-spin  parameters (\eqref{eq:pseudo_spin}) from (\ref{eq:GP}) we obtain
%
%
% 
\begin{equation}
\begin{array}{lcl}
\dot{S}_x & = & -2 \alpha'(t) S_z S_y; \\
\dot{S}_y & = & 2(\alpha'(t) - \beta'(t)) S_z S_x + S_z. \\
\dot{S}_z & = &  2 \beta'(t) S_x S_y - S_y;
\end{array}
\label{eq:pseudo-spin}
\end{equation}
%
Here we explore the dimensionless time $t' = t B / \hbar$.

Set of Eqs.\ ({\ref{eq:pseudo-spin}}) discrebes dynamics of normalized mean field pseudo-spin parameters on the Bloch sphere with $S_x^2 + S_y^2 + S_z^2 = 1$.
Equations  \ ({\ref{eq:pseudo-spin}}) looks similar to one taken for non-dissipative system but with time dependent parameters $\alpha'(t) = (1/ \Lambda) \exp(-2 \kappa' t')$, $\beta'(t) = (\lambda / \Lambda) \exp(-2 \kappa' t')$, $\kappa' = \kappa / B$, cf.\cite{Sedov}. Without dissipation Eqs.\ ({\ref{eq:pseudo-spin}}) posses  bifurcation point $\lambda / \Lambda= 1 / 2$ that corresponds to solid black curve in Fig.1b.

In Fig. \ref{pic:phase}a we plot phase difference $ \phi = \arctan(-S_y / S_x)$ between the junctions as a function on time.  Non-dissipative system has two different regimes depending on the values of  $\beta$ -- parameter, that is blue  (labeled as 1) and red (labeled as 3) curves in the Fig. \ref{pic:phase}, respectively. Blue  curve is plotted under the condition $\beta' > 1/2$ and  corresponds to the quantum regime with existence of barrier for potential energy -- see Fig. \ref{pic:phase}b. Contrary,  red curve establishes phase properties below the threshold ($\beta' <1/2$) that corresponds to the absence of the barrier but upon the same temperature.
